### 2014-06-10 ApTranscriptome_TR.Rmd
Rolling back to analysis with High and Low categories

**June 2, 2014**

@@

```{r max_exp_PDF}
# reshape data
Ap.df <- data.frame(Transcript = rep(Ap.response.type$Transcript, times = 2), colony = rep(c("ApVT", "AcNC"), each = length(Ap.response.type$Transcript)), max.val = c(Ap.response.type$A22.max, Ap.response.type$Ar.max))

maxexplot <- ggplot(Ap.df, aes(x=max.val, fill=colony)) + 
  geom_density(alpha=0.2, position="identity") + 
  #scale_fill_manual(name = "Colony", values = c("white", "black")) + 
  scale_y_continuous(name="Density") +
  scale_x_continuous(name="Temperature of maximum expression")
suppressWarnings(print(maxexplot))
```

@@

```{r optimum_expression_comparison}
# list of transcripts that are 'high' expressed in A22
A22.high.transcripts <- Ap.response.type[which(Ap.response.type$A22.type == "High"), ]

# Compare expression at optimum temp (A22.opt) between colonies using t-test
t.test(A22.high.transcripts$A22.opt, A22.high.transcripts$Ar.opt)
boxplot(A22.high.transcripts$A22.opt, A22.high.transcripts$Ar.opt)

# T test on log-transformed values to control for outliers
t.test(log(A22.high.transcripts$A22.opt+1), log(A22.high.transcripts$Ar.opt+1))
boxplot(log(A22.high.transcripts$A22.opt+1), log(A22.high.transcripts$Ar.opt+1))
```

@@

```{r A22_high_wilcoxon}
w1 <- wilcox.test(A22.high.transcripts$A22.opt, A22.high.transcripts$Ar.opt, alternative = "two.sided", paired = TRUE, conf.int = TRUE)
w1
```

@@

```{r Ar_low_wilcoxon}
# list of transcripts that are 'high' expressed in Ar
Ar.low.transcripts <- Ap.response.type[which(Ap.response.type$Ar.type == "Low"), ]

# t-test with log values
t.test(log(Ar.low.transcripts$A22.opt+1), log(Ar.low.transcripts$Ar.opt+1))
boxplot(log(Ar.low.transcripts$A22.opt+1), log(Ar.low.transcripts$Ar.opt+1))

# Wilcoxon signed rank-test
w2 <- wilcox.test(Ar.low.transcripts$A22.opt, Ar.low.transcripts$Ar.opt, alternative = "two.sided", paired = TRUE, conf.int = TRUE)
w2
```

@@

```{r intermediate_transcripts}
# list of transcripts that are 'Intermediate' expressed in Ar
Ar.int.transcripts <- Ap.response.type[which(Ap.response.type$Ar.type == "Intermediate"), ]
# T test
t.test(log(Ar.int.transcripts$A22.opt+1), log(Ar.int.transcripts$Ar.opt+1))
# Wilcoxon signed rank-test
wilcox.test(Ar.int.transcripts$A22.opt, Ar.int.transcripts$Ar.opt, alternative = "two.sided", paired = TRUE, conf.int = TRUE)

# list of transcripts that are 'Intermediate' expressed in A22
A22.int.transcripts <- Ap.response.type[which(Ap.response.type$A22.type == "Intermediate"), ]
# T test
t.test(log(A22.int.transcripts$A22.opt+1), log(A22.int.transcripts$Ar.opt+1))
# Wilcoxon signed rank-test
wilcox.test(A22.int.transcripts$A22.opt, A22.int.transcripts$Ar.opt, alternative = "two.sided", paired = TRUE, conf.int = TRUE)

# list of transcripts that are 'Intermediate' expressed in either colony
Ap.int.transcripts <- Ap.response.type[which(Ap.response.type$Ar.type == "Intermediate" | Ap.response.type$A22.type == "Intermediate"), ]
# T test
t.test(log(Ap.int.transcripts$A22.opt+1), log(Ap.int.transcripts$Ar.opt+1))
# Wilcoxon signed rank-test
w3 <- wilcox.test(Ap.int.transcripts$A22.opt, Ap.int.transcripts$Ar.opt, alternative = "two.sided", paired = TRUE, conf.int = TRUE)
w3
```

@@

```{r thermal_breadth}
# extract 'Intermediate' expressed transcripts for A22
A22.int.lm <- RxNlmAIC[A22.int.transcripts$Transcript]
length(A22.int.lm)

@@

# repeat for Ar
Ar.int.lm <- RxNlmAIC[Ar.int.transcripts$Transcript]
Ar.int.sd <- unlist(Map(transcriptSD, Ar.int.lm, colony = "Ar"))
Ar.int.sd <- data.frame(colony = "Ar", exp_sd = Ar.int.sd)

@@

```{r extract_Transcripts}
A22.high <- Ap.response.type[which(Ap.response.type$A22.type == "High"), "Transcript"]
Ar.high <- Ap.response.type[which(Ap.response.type$Ar.type == "High"), "Transcript"]

A22.low <- Ap.response.type[which(Ap.response.type$A22.type == "Low"), "Transcript"]
Ar.low <- Ap.response.type[which(Ap.response.type$Ar.type == "Low"), "Transcript"]

A22.bim <- Ap.response.type[which(Ap.response.type$A22.type == "Bimodal"), "Transcript"]
Ar.bim <- Ap.response.type[which(Ap.response.type$Ar.type == "Bimodal"), "Transcript"]

A22.int <- Ap.response.type[which(Ap.response.type$A22.type == "Intermediate"), "Transcript"]
Ar.int <- Ap.response.type[which(Ap.response.type$Ar.type == "Intermediate"), "Transcript"]
```

@@

```{r T_on_high}
# transcripts expressed at *High* and *Bimodal* temperatures in A22
A22.high.TPM.dt.sub <- resp.TPM.dt.sub.pred[J(union(A22.high, A22.bim), "A22")]
setkey(A22.high.TPM.dt.sub, Transcript)
str(A22.high.TPM.dt.sub)

@@

# repeat for Ar
Ar.high.TPM.dt.sub <- resp.TPM.dt.sub.pred[J(union(Ar.high, Ar.bim), "Ar")]
setkey(Ar.high.TPM.dt.sub, Transcript)
str(Ar.high.TPM.dt.sub)

@@

```{r T_on_low}
# transcripts expressed at *Low* temperatures in A22
A22.low.TPM.dt.sub <- resp.TPM.dt.sub.pred[J(union(A22.low, A22.bim), "A22")]
setkey(A22.low.TPM.dt.sub, Transcript)
str(A22.low.TPM.dt.sub)

@@

# repeat for Ar
Ar.low.TPM.dt.sub <- resp.TPM.dt.sub.pred[J(union(Ar.low, Ar.bim), "A22")]
setkey(Ar.low.TPM.dt.sub, Transcript)
str(Ar.low.TPM.dt.sub)

@@

```{r ggplot_high}
trp_A22_high <- ggplot(resp.TPM.dt.sub[A22.high], aes(x=val, y=TPM.scaled, group=Transcript)) + 
  geom_smooth(method = "lm", formula = y ~ poly(x, 2)) + 
  facet_grid(. ~ colony2) + 
  scale_y_continuous(name="Expression (scaled)") +
  scale_x_continuous(name=expression(paste("Temperature ", degree, "C")))
print(trp_A22_high)
```

>>>>

**June 10, 2014**

@@

```{r max_exp_PDF, warning = FALSE, message = FALSE}
# reshape data
Ap.df <- data.frame(Transcript = rep(Ap.response.type$Transcript, times = 2), colony = rep(c("ApVT", "AcNC"), each = length(Ap.response.type$Transcript)), max.val = c(Ap.response.type$A22.max, Ap.response.type$Ar.max))

maxexplot <- ggplot(Ap.df, aes(x=max.val, fill=colony)) + 
  geom_density(alpha=0.2, position="identity") + 
  #scale_fill_manual(name = "Colony", values = c("white", "black")) + 
  scale_y_continuous(name="Density") +
  scale_x_continuous(name="Temperature of maximum expression")
print(maxexplot)
```

@@

```{r optimum_expression_comparison}
# list of transcripts that are 'high' expressed in A22
A22.high <- Ap.response.type[which(Ap.response.type$A22.type == "High"), ]

# Compare expression at optimum temp (A22.opt) between colonies using t-test
t.test(A22.high$A22.opt, A22.high$Ar.opt)
boxplot(A22.high$A22.opt, A22.high$Ar.opt)

# T test on log-transformed values to control for outliers
t.test(log(A22.high$A22.opt+1), log(A22.high$Ar.opt+1))
boxplot(log(A22.high$A22.opt+1), log(A22.high$Ar.opt+1))
```

@@

```{r A22_high_wilcoxon}
w1 <- wilcox.test(A22.high$A22.opt, A22.high$Ar.opt, alternative = "two.sided", paired = TRUE, conf.int = TRUE)
w1
```

@@

```{r Ar_low_wilcoxon}
# list of transcripts that are 'high' expressed in Ar
Ar.low <- Ap.response.type[which(Ap.response.type$Ar.type == "Low"), ]

# t-test with log values
t.test(log(Ar.low$A22.opt+1), log(Ar.low$Ar.opt+1))
boxplot(log(Ar.low$A22.opt+1), log(Ar.low$Ar.opt+1))

# Wilcoxon signed rank-test
w2 <- wilcox.test(Ar.low$A22.opt, Ar.low$Ar.opt, alternative = "two.sided", paired = TRUE, conf.int = TRUE)
w2
```

@@

```{r intermediate_transcripts}
# list of transcripts that are 'Intermediate' expressed in either colony
Ap.int.transcripts <- Ap.response.type[which(Ap.response.type$Ar.type == "Intermediate" | Ap.response.type$A22.type == "Intermediate"), ]
# T test
t.test(log(Ap.int.transcripts$A22.opt+1), log(Ap.int.transcripts$Ar.opt+1))
# Wilcoxon signed rank-test
w3 <- wilcox.test(Ap.int.transcripts$A22.opt, Ap.int.transcripts$Ar.opt, alternative = "two.sided", paired = TRUE, conf.int = TRUE)
w3
```

@@

```{r thermal_breadth}
# extract 'Intermediate' expressed transcripts for A22
A22.int.transcripts <- Ap.response.type[which(Ap.response.type$A22.type == "Intermediate"), ]
A22.int.lm <- RxNlmAIC[A22.int.transcripts$Transcript]
length(A22.int.lm)

@@

# repeat for Ar
Ar.int.transcripts <- Ap.response.type[which(Ap.response.type$Ar.type == "Intermediate"), ]
Ar.int.lm <- RxNlmAIC[Ar.int.transcripts$Transcript]
Ar.int.sd <- unlist(Map(transcriptSD, Ar.int.lm, colony = "Ar"))
Ar.int.sd <- data.frame(colony = "Ar", exp_sd = Ar.int.sd)

@@

```{r extract_Transcripts}
A22.high.transcripts <- Ap.response.type[which(Ap.response.type$A22.type == "High"), "Transcript"]
Ar.high.transcripts <- Ap.response.type[which(Ap.response.type$Ar.type == "High"), "Transcript"]

A22.low.transcripts <- Ap.response.type[which(Ap.response.type$A22.type == "Low"), "Transcript"]
Ar.low.transcripts <- Ap.response.type[which(Ap.response.type$Ar.type == "Low"), "Transcript"]

A22.bim.transcripts <- Ap.response.type[which(Ap.response.type$A22.type == "Bimodal"), "Transcript"]
Ar.bim.transcripts <- Ap.response.type[which(Ap.response.type$Ar.type == "Bimodal"), "Transcript"]

A22.int.transcripts <- Ap.response.type[which(Ap.response.type$A22.type == "Intermediate"), "Transcript"]
Ar.int.transcripts <- Ap.response.type[which(Ap.response.type$Ar.type == "Intermediate"), "Transcript"]
```

@@

```{r T_on_high}
# transcripts expressed at *High* and *Bimodal* temperatures in A22
A22.high.TPM.dt.sub <- resp.TPM.dt.sub.pred[J(union(A22.high.transcripts, A22.bim.transcripts), "A22")]
setkey(A22.high.TPM.dt.sub, Transcript)
str(A22.high.TPM.dt.sub)

@@

# repeat for Ar
Ar.high.TPM.dt.sub <- resp.TPM.dt.sub.pred[J(union(Ar.high.transcripts, Ar.bim.transcripts), "Ar")]
setkey(Ar.high.TPM.dt.sub, Transcript)
str(Ar.high.TPM.dt.sub)

@@

```{r T_on_low}
# transcripts expressed at *Low* temperatures in A22
A22.low.TPM.dt.sub <- resp.TPM.dt.sub.pred[J(union(A22.low.transcripts, A22.bim.transcripts), "A22")]
setkey(A22.low.TPM.dt.sub, Transcript)
str(A22.low.TPM.dt.sub)

@@

# repeat for Ar
Ar.low.TPM.dt.sub <- resp.TPM.dt.sub.pred[J(union(Ar.low.transcripts, Ar.bim.transcripts), "A22")]
setkey(Ar.low.TPM.dt.sub, Transcript)
str(Ar.low.TPM.dt.sub)

@@

```{r ggplot_high}
trp_A22_high <- ggplot(resp.TPM.dt.sub[A22.high.transcripts], aes(x=val, y=TPM.scaled, group=Transcript)) + 
  geom_smooth(method = "lm", formula = y ~ poly(x, 2)) + 
  facet_grid(. ~ colony2) + 
  scale_y_continuous(name="Expression (scaled)") +
  scale_x_continuous(name=expression(paste("Temperature ", degree, "C")))
print(trp_A22_high)
```
